### 2022-03-23 paper3.tex
Final committee edits

This paper examines how social norms are formed under conditions of uncertainty
using the case studies of stay-at-home and vaccination rates during the COVID-19
pandemic. I test fear-based models of behavioral adaption to public health
recommendations as well as patterns of complex social contagion using linear
mixed effects models. These models show that complex contagion is a valid
framework for the social contagion of new norms during COVID-19. However, I find
important moderating effect of signal discordance, the contextual diversity of
signals received by an ego.

% need to explain a little bit more what diversity of signals etc means (since this is the first look) 
% what else needs to be said here to properly convey the importance and impact of this finding?

@@

% TODO
\citep{mirbabaie_etal20, ternullo22}

@@

To make the contagion more complex, different sources of exposure
(county-alters) adhere to CDC recommendations to stay-at-home at differing
rates. Whereas one county-alter may be greatly increasing its time in residence,
another may have made little change. When the majority of alters is in
concordance with each other, the signal to the ego is reinforced and more
impactful on the ego. When these signals are mixed with high variance from
different sources, agreement is low and makes the behavioral change less likely.
For this paper, I theorize that a new concept of `discordance' must be
considered as impacting complex contagion. discordance is the variance of
signals received by an alter; high discordance prevents reinforcement while low
discordance (concordance) enables complex contagion. Instead of adopting a
`majority' rules attitude, this means that the more discordance perceived by an
ego, the less likely the alters will have any effect on the ego. I theorize that
the behavior of a county-alter will be positively correlated with the behavior
of another county-ego if the county-ego receives highly discordant signals, as
seen in figure \ref{fig:dag}.

@@

While much of the social contagion literature, like the theories above, focuses
on structural boundaries and homophily as causes of how diffusion occurs,
\citet{goldbergSocialContagionAssociative2018} propose a
disrupting alternative mechanism. They argue that what actually diffuses during
social contagion are the perceptions about which beliefs or behaviors are
compatible with one another, what they call ``associative diffusion.'' This
argument that culture does not spread like a virus but instead is dependent on
how belief structures are connected to each other is important to test because
norms around health behaviors became politicized issues during the COVID-19
pandemic. This means that the stay-at-home or vaccination behavior themselves
were not contagious, but the cognitive association of precisely what social
distancing or vaccination uptake \emph{meant} were spreading between populations of
individuals. Moreover, as \citet{houghton20} outlines, diffusants - the things that are
diffusing through a population, are not independent of each other \citep{mason_etal07}. When we take into account the interdependence between different beliefs that are diffusing, such as COVID-19 is a hoax, social distancing saves
lives, the COVID-19 vaccine includes a microchip to control your thoughts, and lockdown
is against the rights guaranteed in the US constitution, we can imagine how
these various diffusants become associated together into structures of belief
through schemas of perception \citep{houghton20}. While this cognitive theory of cultural variation is difficult to test, the theory it supplies provides a solid framework for how behavioral norms formed during the pandemic.

The figure above illustrates how I interpret associative diffusion will impact
stay-at-home and vaccination rates in times of unsettled norms. An
individual enters this framework when they realize they don't have a normative
set of responses in their cultural toolkit to respond to an unfamiliar situation
they are presented with. Individuals look to norms to regulate behavior, avoid
deviance, and to maintain order \citep{horneNormsIntegratedFramework2020, shepherdStructurePerceptionHow2017}; when they don't have a normative behavior
to follow (or rebel against), individuals look to the other sources in their
``community'' to mimic behavior, such as high-status individuals, institutions,
and members of their social network. I theorize that individuals look towards
their physical community, to their social network, popular media (which may
include government and science recommendations), established norms that they may
find online through search, and to the real threat of infection (what would
happen if I do nothing about this norm). Following both the Integrated
Theoretical Framework of Norms \citep{horneNormsIntegratedFramework2020} and
associative diffusion \citep{dellapostaWhyLiberalsDrink2015, goldbergSocialContagionAssociative2018}, I theorize that individuals interpret
signals from their ``community'' through their cognitive biases and behavioral
predispositions to determine their formation of new behavioral norms.

There is some evidence I can draw upon to support the associative diffusion
fraemwork. For instance, there is a central conflict between religious and
scientific ideologies which I theorize leads more religious counties to reject
the stay-at-home order and vaccines \citep{evansReligionScienceEpistemological2008}
because the associative diffusion of these behaviors pits public health
recommendations against religious ideologies.
Furthermore, as COVID-19 has become a politically polarizing issue, conservative
ideology and a general mistrust of ``big government'' \citep{frank2007, gauchat2008} are likely to lead to resistance to the
government and scientific guidance.

@@

\begin{figure}
{\centering \includegraphics[width=0.7\linewidth]{figs/paper3/assortativity-vacc-1}}
\caption{How Signal Mean and Signal Assortativity are Measured}\label{fig:assortativity-vacc}
\end{figure}

\begin{figure}
\begin{center}\includegraphics[width=0.5\linewidth,]{figs/paper3/discordancenetwork-1} 
  \begin{equation}
    w = \begin{bmatrix}0.0000015\\0.0000023\\0.0000527\\0.0000966\\0.0001119\end{bmatrix},  
    x = \begin{bmatrix}29.9\\20.4\\27.7\\24.6\\24.3\end{bmatrix} \longrightarrow
    \begin{matrix} \bar{x}, signal = 25.083\\ s, discordance = 1.41 \end{matrix} 
  \end{equation}
  \caption{Visualization of how Network Signal and Discordance are Calculated}
  \label{fig:discordancenetwork}
  \end{center}
\end{figure}

@@

\hypertarget{online-norms}{%
\subsubsection{Online Norms}\label{online-norms}}

% TODO Mention that article-1 says that this can show interest (cite jungherr also) but does not show actions

To operationalize the search for online norms, I use Google search trends \citep{googletrends},
over the study period across individual designated media markets areas (DMAs), a
nonoverlapping aggregation of U.S. counties to 210 media markets based on
similar population clusters \citep{dma_key}. To investigate the rate of \emph{searching} for norms
online in both cases, I use the following search topics: `Social Distancing'
(2020, stay-at-home case only), `COVID-19 Vaccine' (2021, vaccination uptake case
only) and Covid Conspiracy (2020-2021, both cases). Search topics are a more
robust measurement than a single search term: topics are aggregations of the
rates of multiple, highly correlated search terms together into a cohesive
topic. For example, while `Beyoncé,' `Beyonce' and `beyonce knowles' are all
separate search terms, `Beyoncé Knowles' encompasses all of these into a single
search topic. While the data is originally on a scale of 0 to 100, with
100 being the maximum search popularity out of all DMAs, Google Search Trend are
now only available cross-sectionally (a single time period across a geography)
or time-series (a single geo-location across time). To remedy this and build a
longitudinal dataset of each search topic, I follow the method proposed in \citet[p. 5]{park_etal}.
This method involves building a dataset of unscaled
cross-sectional values, selecting a DMA to use to establish the rescaling ratio
(I use `Los Angeles CA'), and then finding the time-series values for the one
DMA. To find the rescaling ratio for each week in the time-series, you divide
the time-series value for each week by the cross-sectional value for each week,
resulting in a rescaling vector to be used for all weeks in the dataset across
geographies. To rescale each longitudinal value, multiply the respective week's
rescaling ratio by the cross-sectional value. Rescaled longitudinal data was
compared against time-series data for multiple test counties and was equivalent.
For a more in-depth explanation of this procedure, see \citet[p. 5]{park_etal}.

@@

My analytic strategy proceeds in four steps. In Tables
\ref{tab:google-desc-table} (Stay-at-Home) and \ref{tab:vacc-desc-table}
(Vaccination Uptake), I present descriptive statistics for all study variables,
including variable ranges, means, and standard deviations across the two cases.
Then, in Tables \ref{tab:google-tab} - \ref{tab:vacc-tab}, I fit a series of
three linear mixed effects regression models using the nlme package in R
\citep{pinheiro_etal21, pinheiro_bates00} for our two cases, Stay-at-Home rates and
Vaccination rates. Linear mixed effects models are a form of hierarchical linear
models that contain both random and fixed effects. These models treat the
dependent variables as continuous. The following strategies I will outline are
identical for both case studies. Models 1 through 3 address hypotheses 1 through
3 respectively. Each model utilizes normalized independent variables, i.e.
variables that have been centered and scaled to have a mean of 0 and standard
deviation 1. The first model is a baseline linear mixed effects model that
employs all basic controls to predict the dependent variable, allowing both
time-varying and county-level variables to predict the outcome. All variables
have fixed effects, meaning that the county-level exogenous effects are
controlled for when estimating the coefficient. In addition, models are
specified with a random intercept per county. The lme models are set with a
autoregressive correlation structure (\texttt{correlation\ =\ corAR1()}) to control for
temporally autocorrelated error structures; models are also optimized using
Nelder--Mead, quasi-Newton and conjugate-gradient algorithms for box-constrained
optimization and simulated annealing (\texttt{control\ =\ lmeControl(opt\ =\ "optim")}).
This first model will address Hypothesis 1. To test hypothesis 2, I estimate
model 2 which builds on model 1 by first introducing vaccination signal. Because
the signal varies by week, I estimate this model with an interaction between
week and signal. And finally, because signal may have divergent effects across
counties, I set a random effect for signal nested by county. A visual
representation of this interaction can be seen in figures
\ref{fig:plot-google-h2} and \ref{fig:plot-vacc-h2}. Model 3 further
elaborates on the previous model by introducing both a fixed effect for signal
discordance and an interaction term between signal and discordance. Figures
\ref{fig:plot-google-h3} and \ref{fig:plot-vacc-h3} depict just how signal and
discordance interact across these two cases.

>>>>

This paper examines how social norms are formed under conditions of uncertainty
using the case studies of stay-at-home and vaccination rates during the COVID-19
pandemic. I test fear-based models of behavioral adaption to public health
recommendations as well as patterns of complex social contagion using linear
mixed effects models. These models show that complex contagion is a valid
framework for the social contagion of new norms during COVID-19. Importantly, I
find a novel moderating effect of signal discordance, the contextual diversity
of signals received by an ego. In other words, if there is diversity in the
information received by an ego, contagion is less likely to occur. This article
is an important contribution to literature on diffusion and social networks, as
it shows that the contagion process cannot be not fully understood without
looking at the context of each exposure to a contagion within  within the range
of contagions one experiences.

@@

@@

To make the contagion more complex, different sources of exposure
(county-alters) adhere to CDC recommendations to stay-at-home at differing
rates. Whereas one county-alter may be greatly increasing its time in residence,
another may have made little change. When the majority of alters is in
concordance with each other, the signal to the ego is reinforced and more
impactful on the ego. When these signals are mixed with high variance from
different sources, agreement is low and makes the behavioral change less likely.
For this paper, I theorize that a new concept of `discordance' must be
considered as impacting complex contagion. discordance is the variance of
signals received by an alter; high discordance prevents reinforcement while low
discordance (concordance) enables complex contagion. Instead of adopting a
`majority' rules attitude, this means that the more discordance perceived by an
ego, the less likely the alters will have any effect on the ego. I theorize that
the behavior of a county-alter will be positively correlated with the behavior
of another county-ego if the county-ego receives highly discordant signals, as
seen in Figure \ref{fig:dag}.

@@

While much of the social contagion literature, like the theories above, focuses
on structural boundaries and homophily as causes of how diffusion occurs,
\citet{goldbergSocialContagionAssociative2018} propose a
disrupting alternative mechanism. They argue that what actually diffuses during
social contagion are the perceptions about which beliefs or behaviors are
compatible with one another, what they call ``associative diffusion.'' This
argument that culture does not spread like a virus but instead is dependent on
how belief structures are connected to each other is important to test because
norms around health behaviors became politicized issues during the COVID-19
pandemic. This means that the stay-at-home or vaccination behavior themselves
were not contagious, but the cognitive association of precisely what social
distancing or vaccination uptake \emph{meant} were spreading between populations of
individuals. Moreover, as \citet{houghton20} outlines, diffusants, the things that are
diffusing through a population, are not independent of each other \citep{mason_etal07}. 
When we take into account the interdependence between different beliefs that are 
diffusing, such as COVID-19 is a hoax, social distancing saves
lives, the COVID-19 vaccine includes a microchip to control your thoughts, and lockdown
is against the rights guaranteed in the US constitution, we can imagine how
these various diffusants become associated together into structures of belief
through schemas of perception \citep{houghton20}. While this cognitive theory of
cultural variation is difficult to test, the theory it supplies provides a solid
framework for how behavioral norms formed during the pandemic.

Individuals look to norms to regulate behavior, avoid
deviance, and to maintain order \citep{horneNormsIntegratedFramework2020, shepherdStructurePerceptionHow2017}. When they realize they don't have a normative
set of responses in their cultural toolkit to respond to an unfamiliar situation
they are presented with, individuals look to the other sources in their
``community'' to mimic behavior, such as high-status individuals, institutions,
and members of their social network. I theorize that individuals look towards
their physical community, to their social network, popular media, established
norms that they may find online through search, and to the real threat of 
infection (what would happen if I do nothing about this norm). 
Following both the Integrated Theoretical Framework of Norms 
\citep{horneNormsIntegratedFramework2020} and
associative diffusion \citep{dellapostaWhyLiberalsDrink2015, goldbergSocialContagionAssociative2018},
I theorize that individuals interpret
signals from their ``community'' through their cognitive biases and behavioral
predispositions to determine their formation of new behavioral norms.

There is some evidence I can draw upon to support the associative diffusion
fraemwork. For instance, there is a central conflict between religious and
scientific ideologies which I theorize leads more religious counties to reject
the stay-at-home order and vaccines \citep{evansReligionScienceEpistemological2008}
because the associative diffusion of these behaviors pits public health
recommendations against religious ideologies.
Furthermore, as COVID-19 has become a politically polarizing issue \citep{ternullo22},
conservative ideology and a general mistrust of ``big government'' \citep{frank2007, gauchat2008} 
are likely to lead to resistance to the government and scientific guidance.

@@

\begin{figure}
{\centering \includegraphics[width=0.7\linewidth]{figs/paper3/assortativity-vacc-1}}
\caption{How Network Signal and Discordance are Calculated, Sampled Regions}\label{fig:assortativity-vacc}
\end{figure}

\begin{figure}
\begin{center}\includegraphics[width=0.5\linewidth,]{figs/paper3/discordancenetwork-1.pdf} 
  \begin{equation}
    w = \begin{bmatrix}0.0000015\\0.0000023\\0.0000527\\0.0000966\\0.0001119\end{bmatrix},  
    x = \begin{bmatrix}29.9\\20.4\\27.7\\24.6\\24.3\end{bmatrix} \longrightarrow
    \begin{matrix} \bar{x}*, signal = 25.083\\ s*, discordance = 1.41 \end{matrix} 
  \end{equation}
  \caption{How Network Signal and Discordance are Calculated, Mathematical Example}
  \label{fig:discordancenetwork}
  \end{center}
\end{figure}

@@

\hypertarget{online-norms}{\subsubsection{Online Norms}\label{online-norms}}

To operationalize the search for online norms, I use Google search trends
\citep{googletrends}. In \hyperlink{paper-1}{\subsection{Article 1}}, I show
that Google Search Trends can only be used to indicate interest due to their low
criterion validity \citep{jungherr_etal17}. I use Google Trends over the study
period across individual designated media markets areas (DMAs), a nonoverlapping
aggregation of U.S. counties to 210 media markets based on similar population
clusters \citep{dma_key}. To investigate the rate of \emph{searching} for norms
online in both cases, I use the following search topics: `Social Distancing'
(2020, stay-at-home case only), `COVID-19 Vaccine' (2021, vaccination uptake
case only) and Covid Conspiracy (2020-2021, both cases). Search topics are a
more robust measurement than a single search term: topics are aggregations of
the rates of multiple, highly correlated search terms together into a cohesive
topic. For example, while `Beyoncé,' `Beyonce' and `beyonce knowles' are all
separate search terms, `Beyoncé Knowles' encompasses all of these into a single
search topic. While the data is originally on a scale of 0 to 100, with 100
being the maximum search popularity out of all DMAs, Google Search Trend are now
only available cross-sectionally (a single time period across a geography) or
time-series (a single geo-location across time). To remedy this and build a
longitudinal dataset of each search topic, I follow the method proposed in
\citet[p. 5]{park_etal}. This method involves building a dataset of unscaled
cross-sectional values, selecting a DMA to use to establish the rescaling ratio
(I use `Los Angeles CA'), and then finding the time-series values for the one
DMA. To find the rescaling ratio for each week in the time-series, you divide
the time-series value for each week by the cross-sectional value for each week,
resulting in a rescaling vector to be used for all weeks in the dataset across
geographies. To rescale each longitudinal value, multiply the respective week's
rescaling ratio by the cross-sectional value. Rescaled longitudinal data was
compared against time-series data for multiple test counties and was equivalent.
For a more in-depth explanation of this procedure, see \citet[p. 5]{park_etal}.

@@

My analytic strategy proceeds in four steps. In Tables
\ref{tab:google-desc-table} (Stay-at-Home) and \ref{tab:vacc-desc-table}
(Vaccination Uptake), I present descriptive statistics for all study variables,
including variable ranges, means, and standard deviations across the two cases.
Then, in Tables \ref{tab:google-tab} - \ref{tab:vacc-tab}, I fit a series of
three linear mixed effects regression models using the nlme package in R
\citep{pinheiro_etal21, pinheiro_bates00} for our two cases, Stay-at-Home rates and
Vaccination rates. Linear mixed effects models are a form of hierarchical linear
models that contain both random and fixed effects. These models treat the
dependent variables as continuous. The following strategies I will outline are
identical for both case studies. Models 1 through 3 address hypotheses 1 through
3 respectively. Each model utilizes normalized independent variables, i.e.
variables that have been centered and scaled to have a mean of 0 and standard
deviation 1. The first model is a baseline linear mixed effects model that
employs all basic controls to predict the dependent variable, allowing both
time-varying and county-level variables to predict the outcome. All variables
have fixed effects, meaning that the county-level exogenous effects are
controlled for when estimating the coefficient. In addition, models are
specified with a random intercept per county. The lme models are set with a
autoregressive correlation structure (\texttt{correlation\ =\ corAR1()}) to control for
temporally autocorrelated error structures; models are also optimized using
Nelder--Mead, quasi-Newton and conjugate-gradient algorithms for box-constrained
optimization and simulated annealing (\texttt{control\ =\ lmeControl(opt\ =\ "optim")}).
This first model will address Hypothesis 1. To test hypothesis 2, I estimate
model 2 which builds on model 1 by first introducing vaccination signal. Because
the signal varies by week, I estimate this model with an interaction between
week and signal. And finally, because signal may have divergent effects across
counties, I set a random effect for signal nested by county. A visual
representation of this interaction can be seen in Figures
\ref{fig:plot-google-h2} and \ref{fig:plot-vacc-h2}. Model 3 further
elaborates on the previous model by introducing both a fixed effect for signal
discordance and an interaction term between signal and discordance. Figures
\ref{fig:plot-google-h3} and \ref{fig:plot-vacc-h3} depict just how signal and
discordance interact across these two cases.

### 2022-03-23 tables/paper1/data-sources-table.tex
3 article summaries
Final committee edits

\begin{table}[!h]
\caption{\label{tab:data-sources-table}Data Sources}
\centering
\fontsize{8}{10}\selectfont
\setlength{\extrarowheight}{2pt}
\begin{tabularx}{\linewidth}{@{}L c c L@{}}
\toprule
\textbf{Validated Data Source} & \textbf{Type} & \textbf{Dates} & \textbf{Google Trends Used}\citep{googletrends} \\ \midrule
% \rowcolor{lightblue}
\multicolumn{4}{l}{\textbf{Behaviors and Attitudes}} \\\hline
% General Social Survey & Cross-Sectional & 2010 - 2020 & NA \\
Vaccine Hesitancy for COVID-19 \citep{vaches_data}& Cross-Sectional & March 3 - 15, 2021 & Search Topics: 'Covid-19 vaccine', 'Coronavirus (Disease)', 'Coronavirus   (Virus)', 'Vaccine' \\
Mask-Wearing Survey Data \citep{mask_data} & Cross-Sectional & July 2 - 14, 2020 & Search Topics: 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Cloth   Face Mask', 'Mask', 'Civil and Political Rights' \\
American National Election Survey \citep{anes_data}& Cross-Sectional & 2020 & NA \\
% \rowcolor{lightblue}
\multicolumn{4}{l}{\textbf{Health}} \\\hline
Covid Rates \citep{covid_data}& Longitudinal & Every Monday, 2020 - 2021 & 'Covid-19', 'Coronavirus', 'Taste Loss', 'Smell Loss' \\
Suicide Rates \citep{suic_data} & Longitudinal & Yearly 2010-2020 & Search Topics: 'Suicide', 'Depression', Search Term:' Suicide Hotline' \\
% \rowcolor{lightblue}
\multicolumn{4}{l}{\textbf{Political}} \\\hline
American National Election Survey \citep{anes_data}& Cross-Sectional & 2020 & NA \\
Presidential Election Results \citep{pres_data}& Cross Sectional & 2016 \& 2020 & Search Topics: 'Hilary Clinton', 'Donald Trump', 'Joe Biden' \\ \bottomrule
\end{tabularx}
\end{table}

>>>>

\begin{table}[!ht]
\caption{\label{tab:data-sources-table}Data Sources}
\centering
\fontsize{8}{10}\selectfont
\setlength{\extrarowheight}{2pt}
\begin{tabularx}{\linewidth}{@{}L c c L@{}}
\toprule
\textbf{Validated Data Source} & \textbf{Type} & \textbf{Dates} & \textbf{Google Trends Used}\citep{googletrends} \\ \midrule
% \rowcolor{lightblue}
\multicolumn{4}{l}{\textbf{Behaviors and Attitudes}} \\\hline
% General Social Survey & Cross-Sectional & 2010 - 2020 & NA \\
Vaccine Hesitancy for COVID-19 \citep{vaches_data}& Cross-Sectional & March 3 - 15, 2021 & Search Topics: 'Covid-19 vaccine', 'Coronavirus (Disease)', 'Coronavirus   (Virus)', 'Vaccine' \\
Mask-Wearing Survey Data \citep{mask_data} & Cross-Sectional & July 2 - 14, 2020 & Search Topics: 'Coronavirus (Disease)', 'Coronavirus (Virus)', 'Cloth   Face Mask', 'Mask', 'Civil and Political Rights' \\
% American National Election Survey \citep{anes_data}& Cross-Sectional & 2020 & NA \\
% \rowcolor{lightblue}
\multicolumn{4}{l}{\textbf{Health}} \\\hline
Covid Rates \citep{covid_data}& Longitudinal & Every Monday, 2020 - 2021 & 'Covid-19', 'Coronavirus', 'Taste Loss', 'Smell Loss' \\
Suicide Rates \citep{suic_data} & Longitudinal & Yearly 2010-2020 & Search Topics: 'Suicide', 'Depression', Search Term:' Suicide Hotline' \\
% \rowcolor{lightblue}
\multicolumn{4}{l}{\textbf{Political}} \\\hline
% American National Election Survey \citep{anes_data}& Cross-Sectional & 2020 & NA \\
Presidential Election Results \citep{pres_data}& Cross Sectional & 2016 \& 2020 & Search Topics: 'Hilary Clinton', 'Donald Trump', 'Joe Biden' \\ \bottomrule
\end{tabularx}
\end{table}
